---
title: "Analysis of Variance: Mixed Model & Family Negative Binomial"
author: "M.A. Hagadorn"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = TRUE,
	background = "#F7F7F7",
	highlight = TRUE)
```

# Load in Required Packages

```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(plyr)
library(dplyr)
library(Rmisc)
library(ggplot2)
library(stats)
library(nortest)
library(MASS)
library(multcomp)
library(effectsize)
library(rcompanion) #plotNormalDensity
library(lme4) #glmer
library(afex)
library(RLRsim)
library(lmerTest)
library(car)  #qqplot
library(lmtest)
library(performance)
library(DHARMa)
library(glmmTMB)
library(reshape2)
library(cowplot)
library(ggbreak)
library(tidyr)
library(emmeans)


installed.packages()[names(sessionInfo()$otherPkgs), "Version"]

#setwd(XXXXX)
```


For this analysis, we are comparing the abundances of dung beetles attracted to specific dung types and habitats. We will do so using a generalized linear mixed model with fixed and random effects. We will compare models of various complexity to determine if the mixed-effects models decrease variance. Further, we are working with count data. So, we will specify "family=Poisson" and then assess overdispersion.





# Load in the Data

To start, we are loading in the data that we plan to analyse, assigning factors and levels. These data were reformatted to long version from the original raw file (SupplementalFile_1.csv)
```{r loaddata}

toab_data0 <- read.csv("SupplementalFile_2.csv", header=T, sep=";")
toab_data <- toab_data0

toab_data$Species <- factor(toab_data$Species)
toab_data$Location <- factor(toab_data$Location)
toab_data$Habitat <- factor(toab_data$Habitat)
toab_data$Type <- factor(toab_data$Type)
toab_data$Animal <- factor(toab_data$Animal)
toab_data$Month <- factor(toab_data$Month, levels = c("June", "July", "Sept"))
```


# Evaluating Models and Model Reduction
## Model 1: Abundance  ~ Habitat * Animal * Month + (1|Location + Species), family = nbinom2
Our initial model includes three fixed and two random effects. Specifically, dung type, habitat where the baited pitfall trap was set, and month are the fixed factors while the species and location (2 different collection sites) are random factors. The first model is as follows: **Abundance  ~ Habitat * Animal * Month + (1|Location + Species)**. 
```{r model1}
mod1 <- glmmTMB(Abundance  ~ Habitat * Animal * Month + (1|Location + Species), data = toab_data, family = poisson) #poisson for count data
#summary(mod1)

#checking for overdispersion before moving forward
check_overdispersion(mod1)

#overdispersion is detected. Evaluating if the negative bionomial model is a better fit.


mod1_negbi <- glmmTMB(Abundance  ~ Habitat * Animal * Month + (1|Location + Species), data = toab_data, family = nbinom2) 

#WARNING: 
#dropping columns from rank-deficient conditional model: Habitatopen:AnimalDeer:MonthJuly, #Habitatopen:AnimalDeer:MonthSept

##compare the two models (family differences) to see which is better, I assume the negbi just correcting for overdispersion


mod_fam <- anova(mod1, mod1_negbi)
mod_fam 
```


The model assuming a poisson distribution shows overdispersion. I compared this model to one assuming a negative bionomial. The model with "family=nbinom2" has a significantly lower AIC score. And so, we will proceed forward using the negative bionomial model.

We did, however, get a warning about rank-deficient conditional model and dropping a single column (Abundance  ~ Habitat * Animal * Month + (1|Location + Species), family = nbinom2). So, before moving forward, I am going to see if the model is better with or without the interaction term. First, I will run summary() on model one to get a gauge of the output.

### Assess the use of both random variables
Now, I am assessing the negative bionomial module 1 using the "summary()" function. 
```{r assessrandomvarVar}
summary(mod1_negbi)

```

Based on the summary of the model, neither "location" or "species" has a variance close to zero, though location is substantially less than species.

## Generate all models to compare best fit
Now, lets generate all combinations of the modules to assess for best fit, starting with more complex to simplified.
```{r models}
mod2_negbi <- glmmTMB(Abundance ~ Habitat + Animal + Month + (1|Location + Species), data = toab_data, family = nbinom2)

mod3_negbi <- glmmTMB(Abundance ~ Habitat + Animal + (1|Location + Species + Month), data = toab_data, family = nbinom2)

mod4_negbi <- glmmTMB(Abundance ~ Habitat + Animal + (1|Location + Species), data = toab_data, family = nbinom2)

mod5_negbi <- glmmTMB(Abundance ~ Habitat + Animal + Month + (1|Species), data = toab_data, family = nbinom2)

mod6_negbi <- glmmTMB(Abundance ~ Habitat + Animal + Month + (1|Location), data = toab_data, family = nbinom2)

mod7_negbi <- glmmTMB(Abundance ~ Habitat * Animal + Month + (1|Location + Species), data = toab_data, family = nbinom2)

mod8_negbi <- glmmTMB(Abundance ~ Habitat * Type * Month + (1|Location + Species), data = toab_data, family = nbinom2)

mod9_negbi <- glmmTMB(Abundance ~ Habitat + Type + Month + (1|Location + Species), data = toab_data, family = nbinom2)

```

## Model 2 comparison: remove interaction term in full model to assess best fit; **Abundance ~ Habitat + Animal + Month + (1|Location + Species)**
Now we compare the negative binomial versions of model 1 and model 2. From this, we can assess if the models are significantly different, i.e., does the interaction term contribute significantly to the model.
```{r comparemods}
modcomp1 <- anova(mod2_negbi, mod1_negbi)
modcomp1

#AIC Difference
abs(modcomp1$AIC[1]-modcomp1$AIC[2])

#P
modcomp1P <- modcomp1$`Pr(>Chisq)`[2]

```

AIC value difference is greater than 2.0 (`r abs(modcomp1$AIC[1] - modcomp1$AIC[2])`) and the P-value comparing the model is significant (p = `r modcomp1P`). That said, given the rank-deficiency of the full model (interaction terms) we are going to proceed forward with other model comparisons.

## Model 3 comparisons: Month as a random factor **Abundance ~ Habitat + Animal + (1|Location + Species + Month)**
### Type of variable assignment of month: fixed or random?
```{r model3}
#compare the model with different placement of month
modcomp2 <- anova(mod3_negbi, mod2_negbi)
modcomp2

#AIC Difference
abs(modcomp2$AIC[1]-modcomp2$AIC[2])

#P
modcomp2P <- modcomp2$`Pr(>Chisq)`[2]

```

The AIC score difference *IS* greater than 2.0 (`r abs(modcomp2$AIC[1] - modcomp2$AIC[2])`) and the P-value comparing the model is significant (p=`r modcomp2P `). The AIC score for model 2 (full model, no interaction, month as fixed factor) is: `r modcomp2$AIC[2]`), whereas the AIC score for model that shifts month to a random factor is: `r modcomp2$AIC[1]`. Thus, the model including month as a fixed factor is a better fit.


## Model 2 (best fit so far) vs reduced models:
### Is month important?
Now, we are comparing the model with month as a fixed factor to a model that excludes month entirely.
```{r model2V4}
#compare the models when one lacks month entirely
modcomp3 <- anova(mod4_negbi, mod2_negbi)
modcomp3

#AIC Difference
abs(modcomp3$AIC[1]-modcomp3$AIC[2])

#P
modcomp3P <- modcomp3$`Pr(>Chisq)`[2]

```

The AIC score difference *IS* greater than 2.0 (`r abs(modcomp3$AIC[1] - modcomp3$AIC[2])`) and the P-value comparing the model is significant (p=`r modcomp3P `). The AIC score for model 2 (full model, no interaction, month as fixed factor) is: `r modcomp3$AIC[2]`), whereas the AIC score for model that shifts month to a random factor is: `r modcomp3$AIC[1]`. Thus, the reduced model is a worse fit. We will proceed forward with the full model: **Abundance ~ Habitat + Animal + Month + (1|Location + Species), data = toab_data, family = nbinom2**.


### Is location important?
Now, we are going to determine if Location is an important factor.
```{r model2v5}
#with location removed
modcomp4 <- anova(mod5_negbi, mod2_negbi)
modcomp4

#AIC Difference
abs(modcomp4$AIC[1]-modcomp4$AIC[2])

#P
modcomp4P <- modcomp4$`Pr(>Chisq)`[2]

```

The AIC score difference *IS* greater than 2.0 (`r abs(modcomp4$AIC[1] - modcomp4$AIC[2])`) and the P-value comparing the model is significant (p=`r modcomp4P `). The AIC score for model 2 (full model, no interaction, month as fixed factor) is: `r modcomp4$AIC[2]`), whereas the AIC score for model that removes location is: `r modcomp4$AIC[1]`. Thus, the reduced model is a worse fit. We will proceed forward with the full model: **Abundance ~ Habitat + Animal + Month + (1|Location + Species), data = toab_data, family = nbinom2**.


### Is species important?
Now, we are going to determine if species is an important factor.
```{r model2v6}
#with location removed
modcomp5 <- anova(mod6_negbi, mod2_negbi)
modcomp5

#AIC Difference
abs(modcomp5$AIC[1]-modcomp5$AIC[2])

#P
modcomp5P <- modcomp5$`Pr(>Chisq)`[2]

```

The AIC score difference *IS* greater than 2.0 (`r abs(modcomp5$AIC[1] - modcomp5$AIC[2])`) and the P-value comparing the model is significant (p=`r modcomp5P `). The AIC score for model 2 (full model, no interaction, month as fixed factor) is: `r modcomp5$AIC[2]`), whereas the AIC score for model excludes species: `r modcomp5$AIC[1]`. Thus, the reduced model is a worse fit. We will proceed forward with the full model: **Abundance ~ Habitat + Animal + Month + (1|Location + Species), data = toab_data, family = nbinom2**.



## Summarize model fit
```{r kablefit}
Models <- c("Abundance  ~ Habitat * Animal * Month + (Loc + Sp)", "Abundance  ~ Habitat * Animal * Month + (Loc + Sp)", "Abundance ~ Habitat + Animal + Month + (Loc + Sp)", "Abundance ~ Habitat + Animal + (Loc + Sp + Month)", "Abundance ~ Habitat + Animal + (Loc + Sp)", "Abundance ~ Habitat + Animal + Month + (Sp)", "Abundance ~ Habitat + Animal + Month + (Loc)")
Family <- c("P", rep("NB", 6))
DFs <- c(mod_fam$Df[1], mod_fam$Df[2], modcomp2$Df[2], modcomp2$Df[1], modcomp3$Df[1], modcomp4$Df[1], modcomp5$Df[1])
AICs <- c(AIC(mod1), AIC(mod1_negbi), AIC(mod2_negbi), AIC(mod3_negbi), AIC(mod4_negbi), AIC(mod5_negbi), AIC(mod6_negbi))
Selected <- c("no", "no", "yes", "no", "no", "no", "no")

sum_tab <- data.frame(Models, Family, DFs, AICs, Selected)


kable(sum_tab)

```







# Run Analysis with full model but no interaction: Abundance  ~ Habitat + Type + (1|Location + Species); family=negative bionomial
For the analysis, we will move forward with the GLMM that incorporates habitat, animal, and month as fixed factors, without testing an interaction, and location and species as a random factor.
```{r assignthefinalmodel}
#this is mod2_negbi
toab_glmm <- glmmTMB(Abundance ~ Habitat + Animal + Month + (1|Location + Species), data = toab_data, family = nbinom2)
```


## Summary Statistics
```{r ss_toab, echo=FALSE}
ss <- summarySE(toab_data, measurevar = "Abundance", groupvars = c("Habitat", "Animal", "Month"))
kable(ss)

# ss_species <- summarySE(toab_data, measurevar = "Abundance", groupvars = c("Species", "Animal", "Month"))
# kable(ss_species)

ss_justAnMonth <- summarySE(toab_data, measurevar = "Abundance", groupvars = c("Animal", "Month"))
kable(ss_justAnMonth)

ss_sponly <- summarySE(toab_data, measurevar = "Abundance", groupvars = c("Species", "Month"))
kable(ss_sponly)

ss_justanimal <- summarySE(toab_data, measurevar = "Abundance", groupvars = c("Animal"))
kable(ss_justanimal)

ss_MT <- summarySE(toab_data, measurevar = "Abundance", groupvars = c("Month", "Type"))
kable(ss_MT)

```

\newpage

```{r ss_toab2, echo=FALSE}
ss_sponly <- summarySE(toab_data, measurevar = "Abundance", groupvars = c("Species", "Month"))
kable(ss_sponly)

ss_justanimal <- summarySE(toab_data, measurevar = "Abundance", groupvars = c("Animal"))
kable(ss_justanimal)

```

## Run Model
```{r glmm_run}
summary(toab_glmm)

#get an anova table
#type two because there is no significant interaction term
Anova(toab_glmm, type = "II", test.statistic = "Chisq")


# Calculate percent more
#percent_more <- function(higher, lower)((higher - lower) / lower) * 100


ratio_pmore <- function(em_df, col, higher, lower) {
  # Ensure the grouping column is character
  em_df[[col]] <- as.character(em_df[[col]])
  
  # Extract the estimated means
  mean_higher <- em_df$response[em_df[[col]] == higher]
  mean_lower  <- em_df$response[em_df[[col]] == lower]
  
  # Calculate ratio and percent more
  ratio <- mean_higher / mean_lower
  percent_more <- (ratio - 1) * 100
  
  # Return results
  list(
    higher = higher,
    lower = lower,
    ratio = ratio,
    percent_more = percent_more
  )
}


##POSTHOC

#Habitat
##emmeans
emm_hab <- emmeans(toab_glmm, ~ Habitat, type = "response", weights = "proportional")
#estimated marginal means
pairs(emm_hab, adjust = "bonferroni")
#Letters
multcomp::cld(emm_hab, adjust = "bonferroni", Letters = letters)

emm_hab_df <- as.data.frame(emm_hab)
emm_hab_df


ratio_pmore(emm_hab_df, "Habitat", "open", "forest")


# percent_more(emm_hab_df$response[emm_hab_df$Habitat == "open"], emm_hab_df$response[emm_hab_df$Habitat == "forest"])
# 
# 
# contrast(emm_hab_df, "ratio", ref = "forest", adjust = "none")


##Type of animal generating dung
##emmeans
emm_animal <- emmeans(toab_glmm, ~ Animal, type = "response", weights = "proportional")
#estimated marginal means
pairs(emm_animal, adjust = "bonferroni")
#Letters
multcomp::cld(emm_animal, adjust = "bonferroni", Letters = letters)

emm_animal_df <- as.data.frame(emm_animal)
emm_animal_df

##bear v wolf
ratio_pmore(emm_animal_df, "Animal", "Bear/Dog", "Wolf")

# percent_more(emm_animal_df$response[emm_animal_df$Animal == "Bear"], emm_animal_df$response[emm_animal_df$Animal == "Wolf"])
# 
# contrast(emm_animal_df, "ratio", ref = "Wolf", adjust = "none")

##bear v deer
ratio_pmore(emm_animal_df, "Animal", "Bear/Dog", "Deer")

# percent_more(emm_animal_df$response[emm_animal_df$Animal == "Bear"], emm_animal_df$response[emm_animal_df$Animal == "Deer"])
# 
# contrast(emm_animal_df, "ratio", ref = "Deer", adjust = "none")

##alpaca v deer
ratio_pmore(emm_animal_df, "Animal", "Alpaca", "Deer")

# percent_more(emm_animal_df$response[emm_animal_df$Animal == "Alpaca"], emm_animal_df$response[emm_animal_df$Animal == "Deer"])



##Month
##emmeans
emm_month <- emmeans(toab_glmm, ~ Month, type = "response", weights = "proportional")
#estimated marginal means
pairs(emm_month, adjust = "bonferroni")
#Letters
multcomp::cld(emm_month, adjust = "bonferroni", Letters = letters)

emm_month_df <- as.data.frame(emm_month)
emm_month_df

##June v July
ratio_pmore(emm_month_df, "Month", "June", "July")


# percent_more(emm_month_df$response[emm_month_df$Month == "June"], emm_month_df$response[emm_month_df$Month == "July"])
# 
# contrast(emm_month_df, "ratio", ref = "July", adjust = "none")

#June v September
ratio_pmore(emm_month_df, "Month", "June", "Sept")
# percent_more(emm_month_df$response[emm_month_df$Month == "June"], emm_month_df$response[emm_month_df$Month == "Sept"])
# 
# contrast(emm_month_df, "ratio", ref = "Sept", adjust = "none")

```


## Testing Parametric Assumptions
```{r assumptions}

res.toab <- residuals(toab_glmm)
par(mfrow=c(1,2))
plotNormalDensity(res.toab)
qqPlot(res.toab, "norm")

#Anderson Darling test for normality
#Ho: The residuals are normally distributed
#Ha: The residuals are not normally distributed
#aka if the p-value is <0.05 then you can conclude that the data are not normally distributed
print(shapiro.test(res.toab))

#Homogeneity of variance
#Ho: The distributions are the same.
#Ha: The distributions are not the same.
#aka if the p-value is <0.05 then you can conclude a lack of homogenetity of variance.

#do levene's test
#levene.test or leveneTest in R can't work with a random variable and a non fully crossed design (i.e., it has to consider that interaction term, even if it isn't in our model.)
leveneTest(residuals(toab_glmm)  ~ Habitat * Animal * Month, data = toab_data)

#another check
check_homogeneity(toab_glmm)


##Assessing the residuals Using DPharm
#https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html

simres_toabglmm <- simulateResiduals(fittedModel = toab_glmm, plot = F)
#This calculates randomized quantile residuals. Contains simulations and the scaled residuals

res_norm <- residuals(simres_toabglmm , quantileFunction = qnorm, outlierValues = c(-7,7))

plotQQunif(simres_toabglmm) # left plot in plot.DHARMa()
plotResiduals(simres_toabglmm) # right plot in plot.DHARMa()

```


```{r checkmodel, fig.height=9, fig.width=7}
#visual
check_model(toab_glmm)
```





# Figures
I'm going to make the figures as DLP and I discussed.

## Figure 1: bar and line
```{r makegraphs}
#get data set up for box plot and line graph
bp_data <- toab_data[,c("Habitat", "Type", "Abundance")]

melt_toab <- melt(bp_data)

#assign order levels
melt_toab$Habitat <- factor(melt_toab$Habitat, levels=c("forest", "open"))
melt_toab$Type <- factor(melt_toab$Type, levels=c("Carnivore", "Herbivore", "Omnivore"))

#Assigning specific sizes, changes to figures will reflect changes made here. Keep consistent with Exp.1 figures
generalsize <- 14
legendsize <- 7
wordsize <- 12
lettersize <- 3
dotsize <- 3
```


```{r boxplots}
# Recalculate summary with Habitat on x-axis
melt_df_plot <- melt_toab %>%
  group_by(Habitat, Type) %>%
  summarise(mean = mean(value), 
            se = sd(value) / sqrt(n()), 
            .groups = "drop")

h_b <- 55
h_t <- 60


# Plot with Habitat on x-axis and bars colored by Type
bp_A <- ggplot(melt_df_plot, aes(x = Habitat, y = mean, fill = factor(Type))) +
  geom_col(position = position_dodge(width = .75), width = 0.9, color = "black") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                position = position_dodge(width = 0.9), width = 0.2) +
  geom_jitter(data = melt_toab, aes(x = Habitat, y = value, color = factor(Type)),
              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.55, seed = 2025),
              alpha = 0.65, size = dotsize) +
  scale_fill_manual(values = c("Carnivore" = "white", "Herbivore" = "gray", "Omnivore" = "gray45")) +  # Customize as needed
  scale_color_manual(values = c("Carnivore" = "darkred", "Herbivore" = "forestgreen", "Omnivore" = "goldenrod2")) +
  scale_y_continuous(name = "Abundance", limits = c(-1, 70), breaks = seq(0, 70, 10), expand = expansion(mult = c(0, 0.01))) +
  scale_x_discrete(name = "", labels = c("Forest", "Open")) +
  ggtitle("A") +
  
  annotate(geom="text", x=1.5, y=h_t + 2, label="*", color="black", size = 10) +
  
  # Bracket
  geom_segment(aes(x = 1, xend = 2, y = h_t, yend = h_t)) +  # horizontal top
  geom_segment(aes(x = 1, xend = 1, y = h_b, yend = h_t)) +  # left vertical
  geom_segment(aes(x = 2, xend = 2, y = h_b, yend = h_t)) +      # right vertical
  
  
  
  theme(
    text = element_text(color = "black", size = generalsize, family = "sans"),
    axis.title = element_text(color = "black", size = 14),
    axis.text.x = element_text(color = "black", size = wordsize),
    axis.text.y = element_text(color = "black", size = wordsize, margin = margin(l = 2, r = 2)),
    axis.ticks.length = unit(0.2, "cm"),
    axis.ticks = element_line(color = "black", size = 0.3),
    legend.position = "none",
    panel.background = element_rect(fill = "white"),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black"),
    plot.margin = margin(t = 1, 0, b = -12, 0),
    plot.title = element_text(size = 16, margin = margin(t = 1, r = 0, b = 0, l = 0), face = "bold", vjust = 1.75, hjust = -.12)
  )

```


```{r linebymonth}

line_data <- toab_data[, c("Type", "Month", "Abundance")]
melt_toab_line <- melt(line_data)

# Assign month order (incl. August even though it's missing from data)
month_levels <- c("June", "July", "August", "Sept")

# Summarize mean and SE by Month and Type (no Habitat)
melt_df_lineavg <- melt_toab_line %>%
  group_by(Month, Type) %>%
  summarise(
    mean = mean(value, na.rm = TRUE),
    se = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  mutate(Month = factor(Month, levels = month_levels, ordered = TRUE))

# Interpolate August between July and September, per Type
august_data <- melt_df_lineavg %>%
  filter(Month %in% c("July", "Sept")) %>%
  group_by(Type) %>%
  summarise(
    mean = mean(mean, na.rm = TRUE),
    se = mean(se, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(Month = factor("August", levels = month_levels, ordered = TRUE))

# Combine original data and artificial August
combined_df <- bind_rows(melt_df_lineavg, august_data)

# Refactor Month after combination and ensure all levels present
combined_df <- combined_df %>%
  mutate(Month = factor(Month, levels = month_levels, ordered = TRUE)) %>%
  arrange(Type, Month)

# Ensure every Month-Type combination exists (will fill with NAs if missing)
melt_df_lineavg_aug <- complete(combined_df, Month, Type)

# Get numeric position of "August" for vertical dotted line
august_pos <- which(levels(melt_df_lineavg_aug$Month) == "August")

# Plot
line_b <- ggplot(melt_df_lineavg_aug, aes(x = Month, y = mean, group = Type, color = Type)) +
  geom_vline(xintercept = august_pos, linetype = "longdash", color = "grey", linewidth = 1.5) +
  geom_line(linewidth = 1.75, na.rm = TRUE) +
  geom_point(data = filter(melt_df_lineavg_aug, !is.na(mean) & Month != "August"), size = 4) +
  geom_errorbar(data = filter(melt_df_lineavg_aug, !is.na(mean) & Month != "August"),
                aes(ymin = mean - se, ymax = mean + se),
                width = 0.2, linewidth = 1.75) +
  scale_color_manual(values = c("Carnivore" = "darkred", "Herbivore" = "forestgreen", "Omnivore" = "goldenrod2"), guide = "none") +
  scale_y_continuous(name = "Abundance", limits = c(-.2, 15), breaks = seq(0, 15, 5),  expand = expansion(mult = c(0, 0.01))) +
  scale_x_discrete(name = NULL, labels = c("June", "July", "August", "September")) +
  #labs(x = "Month") +
    ggtitle("B") +
  
  annotate(geom="text", x=1, y=14.5, label="a", color="black", size = 6) +
  annotate(geom="text", x=2, y=14.5, label="b", color="black", size = 6) + 
  annotate(geom="text", x=4, y=14.5, label="b", color="black", size = 6) +
  
  
theme(text = element_text(color = "black", size = generalsize, family = "sans"),
    axis.title = element_text(color = "black", size =14),
    axis.text.x = element_text(color = "black", size = wordsize),  # simplified
    axis.text.y = element_text(color = "black", size = wordsize, margin = margin(l = 2, r = 2)),
    axis.ticks.margin = unit(0, "cm"),
    axis.ticks.length = unit(0.2, "cm"),
    axis.ticks = element_line(color = "black", size = 0.3),
    #legend.title = element_blank(),
    legend.position = "none",
    #legend.margin = margin(l = 0.6, r = 0.6, t = 0.5, b = 0.5),
    #legend.spacing.y = unit(0, "pt"),
    #legend.spacing.x = unit(0.3, "pt"),
    #legend.background = element_rect(fill = "white", size = 0.25, linetype = "solid", colour = "black"),
    #legend.key = element_rect(fill = "white"),
    #legend.text = element_text(size = legendsize),
    #legend.key.height = unit(17, "pt"),
    #legend.key.width = unit(0.4, "cm"),
    panel.background = element_rect(fill = "white"),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black"),
    plot.margin = margin(t = 1, 0, b = 4.5, 0),
    plot.title = element_text(size = 16, margin = margin(t = 1, r = 0, b = 0, l = 0), face = "bold", vjust = 1.75, hjust = -.12))
```


```{r panel}
fig1 <- plot_grid(bp_A, line_b, 
          #labels = c("A", "B"), 
          #label_size = 16, 
          ncol = 2, 
          align = "v")  # vertical alignment of axes


fig1

ggsave("Fig1_panel.tiff", plot = fig1, device = "tiff", 
       width = 10, height = 5, units = "in", dpi = 600) 

ggsave("Fig1_panel.png", plot = fig1, device = "png", 
       width = 10, height = 5, units = "in", dpi = 600) 

```



## Figure 2: Animal dung, species, and abundance
```{r bigboyplot, fig.height=9, fig.width=8}
library(dplyr)
library(tidyr)
library(ggplot2)


# Prepare and aggregate data by Species–Animal pair
fig2_data <- toab_data %>%
  dplyr::group_by(Species, Animal) %>%
  dplyr::summarise(Abundance = sum(Abundance), .groups = "drop")



# Normalize within species and animal
fig2_data <- fig2_data %>%
  dplyr::group_by(Species) %>%
  dplyr::mutate(Prop_Species = Abundance / sum(Abundance)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Animal) %>%
  dplyr::mutate(Prop_Animal = Abundance / sum(Abundance)) %>%
  dplyr::ungroup()

# Total abundance proportions for species and animals
species_totals <- fig2_data %>%
  dplyr::group_by(Species) %>%
  dplyr::summarize(Total_Abundance = sum(Abundance)) %>%
  dplyr::mutate(Proportion = Total_Abundance / sum(Total_Abundance)) %>%
  dplyr::arrange(desc(Proportion))

animal_totals <- fig2_data %>%
  dplyr::group_by(Animal) %>%
  dplyr::summarize(Total_Abundance = sum(Abundance)) %>%
  dplyr::mutate(Proportion = Total_Abundance / sum(Total_Abundance)) %>%
  dplyr::arrange(desc(Proportion))

# Add gaps between bars
gap_species <- 1 / (nrow(species_totals) * 5)
gap_animal <- 1 / (nrow(animal_totals) * 5)

species_heights <- species_totals %>%
  dplyr::mutate(
    ymin = cumsum(dplyr::lag(Proportion + gap_species, default = 0)),
    ymax = ymin + Proportion
  )

animal_heights <- animal_totals %>%
  dplyr::mutate(
    ymin = cumsum(dplyr::lag(Proportion + gap_animal, default = 0)),
    ymax = ymin + Proportion
  )

# Join coordinates to main data
fig2_data <- fig2_data %>%
  dplyr::left_join(species_heights, by = "Species") %>%
  dplyr::rename(S_ymin = ymin, S_ymax = ymax) %>%
  dplyr::left_join(animal_heights, by = "Animal") %>%
  dplyr::rename(A_ymin = ymin, A_ymax = ymax)

# Compute link positions within bars
fig2_data <- fig2_data %>%
  dplyr::group_by(Species) %>%
  dplyr::mutate(
    S_total = sum(Abundance),
    S_prop = Abundance / S_total,
    S_ymin_link = dplyr::lag(cumsum(S_prop * (S_ymax - S_ymin)), default = 0) + S_ymin,
    S_ymax_link = S_ymin_link + S_prop * (S_ymax - S_ymin)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Animal) %>%
  dplyr::mutate(
    A_total = sum(Abundance),
    A_prop = Abundance / A_total,
    A_ymin_link = dplyr::lag(cumsum(A_prop * (A_ymax - A_ymin)), default = 0) + A_ymin,
    A_ymax_link = A_ymin_link + A_prop * (A_ymax - A_ymin)
  ) %>%
  dplyr::ungroup()

# Assign plotting priority and sort ribbons: highest priority (largest) last so they overlay
fig2_data <- fig2_data %>%
  dplyr::mutate(
    ribbon_prop = pmax(S_prop, A_prop),
    priority = dplyr::case_when(
      (Animal == "Bear") ~ 3L,                                # Highest priority: drawn last (top)
      (Animal == "Alpaca" & Species == "oscarinus") ~ 2L,    # Next highest
      Species %in% c("Omorgus", "Alloblackburneous") ~ 0L,    # Lowest priority: drawn first (bottom)
      TRUE ~ 1L
    )
  ) %>%
  dplyr::arrange(priority, desc(ribbon_prop))  # sort so low priority drawn first

# Build polygon coordinates for ribbons
polygon_data <- fig2_data %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    x = list(c(1.5, 4.5, 4.5, 1.5)),
    y = list(c(S_ymin_link, A_ymin_link, A_ymax_link, S_ymax_link))
  ) %>%
  tidyr::unnest(c(x, y)) %>%
  dplyr::group_by(Species, Animal, priority) %>%
  dplyr::mutate(point_id = dplyr::row_number()) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(priority, Species, Animal, point_id)

# Define unique colors for each Animal
animal_colors <- c(
  "Bear/Dog" = "darkgoldenrod2",
  "Deer" = "#009E73",
  "Alpaca" = "#56B4E9",
  "Wolf" = "blueviolet",
  "Bobcat" = "orangered"
)

#species order with correct naming
Species_aes <- as.character(species_heights$Species)
Species_aes <- gsub("_", " ", Species_aes)
Species_aes <- sub("s1", "s 1", Species_aes)
Species_label <- dplyr::case_when(
  Species_aes == "Trox species 1" ~ "italic('Trox')~'species 1'",
  Species_aes == "Copris minitus" ~ "italic('Copris minutus')", #correct spelling
  Species_aes == "Onthophagus pennsyvanicus" ~ "italic('Onthophagus pennsylvanicus')", #correct spelling
  TRUE ~ paste0("italic('", Species_aes, "')")  # default
)# Species_label <- dplyr::case_when(
#   Species_aes == "Trox species 1" ~ "italic('Trox')~'species 1'",
#   TRUE ~ paste0("italic('", Species_aes, "')")
# )






# Plot
fig2 <- ggplot() +
  # Species bars
  geom_rect(data = species_heights,
            aes(xmin = 1, xmax = 1.5, ymin = ymin, ymax = ymax),
            fill = "gray45") +
  # Animal bars
  geom_rect(data = animal_heights,
            aes(xmin = 4.5, xmax = 5, ymin = ymin, ymax = ymax),
            fill = "gray45") +
  # Connecting ribbons colored by Animal
  geom_polygon(
    data = polygon_data,
    aes(x = x, y = y, group = interaction(Species, Animal), fill = Animal, alpha=0.75),
    color = "black",
    linewidth = 0.3
  ) +
  # Species labels
  geom_text(data = species_heights,
            aes(x = .9, y = (ymin + ymax) / 2, label = Species_label),
            hjust = 1, size = 4, parse = TRUE, family = "sans") +
  
  # Animal labels
  geom_text(data = animal_heights,
            aes(x = 5.1, y = (ymin + ymax) / 2, label = Animal),
            hjust = 0, size = 4, family = "sans") +
  
  scale_fill_manual(values = animal_colors) +
  coord_cartesian(clip = "off") +
  theme_void() +
  theme(
    plot.margin = margin(t=5, r=49, b=2, l=149, unit = "pt"),
    #aspect.ratio = 1.3,
    legend.position = "none", 
    #plot.background = element_rect(fill = "lightblue"),
    #panel.background = element_rect(fill = "pink")
  ) +

  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0))

fig2

ggsave("Fig2.tiff", plot = fig2, device = "tiff", 
      width = 8, height = 10, units = "in", dpi = 1000, bg = "white") 

ggsave("Fig2.png", plot = fig2, device = "png", 
       width = 8, height = 10, units = "in", dpi = 1000, bg = "white") 

```

